---
title: "UFC Statistics"
author: "Team #5"
date: "7/24/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(car)
library(dplyr)
library(ggplot2)
library(reshape)
library(pastecs)
library(gmodels)
# library(here)
```

Link to the data set: https://www.kaggle.com/mdabbert/ultimate-ufc-dataset?select=ufc-master.csv

```{r load_data}
df <- read.csv('ufc-master.csv')

# head(df)
# names(df)
# summary(df)
```

# Inquiry 1: Does one color have an advantage over the other? Does gender make a difference?

```{r color_analysis_1}

# Get the data
color <- df[,c('Winner', 'gender')]

# Plot the graph without gender
colorOnly.bar <- ggplot(color, aes(x = Winner))
colorOnly.bar +
  geom_bar(fill = c('Blue', 'Red'))

```

Based on the graph, it appears that red has a large advantage over blue. (Question: How are colors chosen? At random? Based on rank?)

We can break this down and look at the gender splits.

```{r color_analysis_2}
# Plot the graph with gender
colorGender.table <- with(color, table(Winner, gender))
colorGender.bar <- barplot(colorGender.table, beside = TRUE, legend = TRUE)

# Calculate win percentages
print(paste('Overall Red Wins: ',
            nrow(subset(color, Winner == 'Red'))))
print(paste('Overall Blue Wins: ',
            nrow(subset(color, Winner == 'Blue'))))

print(paste('Overall Red Win Pct: ',
            round( 100 * nrow(subset(color, Winner == 'Red')) / nrow(color), 2 )))
print(paste('Overall Blue Win Pct: ',
            round( 100 * nrow(subset(color, Winner == 'Blue')) / nrow(color), 2 )))

print(paste('Overall Male Red Wins: ',
            nrow(subset(color, Winner == 'Red' & gender == 'MALE'))))
print(paste('Overall Male Blue Wins: ',
            nrow(subset(color, Winner == 'Blue' & gender == 'MALE'))))

print(paste('Overall Male Red Win Pct: ',
            round( 100 * nrow(subset(color, Winner == 'Red' & gender == 'FEMALE')) /
                     nrow(subset(color, gender == 'FEMALE')), 2)))
print(paste('Overall Male Blue Win Pct: ',
            round( 100 * nrow(subset(color, Winner == 'Blue' & gender == 'FEMALE')) /
                     nrow(subset(color, gender == 'FEMALE')), 2 )))

print(paste('Overall Female Red Wins: ',
            nrow(subset(color, Winner == 'Red' & gender == 'FEMALE'))))
print(paste('Overall Female Blue Wins: ',
            nrow(subset(color, Winner == 'Blue' & gender == 'FEMALE'))))

print(paste('Overall Female Red Win Pct: ',
            round( 100 * nrow(subset(color, Winner == 'Red' & gender == 'FEMALE')) /
                     nrow(subset(color, gender == 'FEMALE')), 2 )))
print(paste('Overall Female Blue Win Pct: ',
            round( 100 * nrow(subset(color, Winner == 'Blue' & gender == 'FEMALE')) /
                     nrow(subset(color, gender == 'FEMALE')), 2 )))

```

I'm not sure what test to run here to show that color is significant.

# Inquiry 2: Does a reach advantage lead to more wins?

```{r reach_analysis}
# Create a smaller data frame
reach <- df[,c('Winner', 'gender', 'B_Reach_cms', 'R_Reach_cms')]

# Calculate the difference in reach (positive = blue advantage)
reach$Diff <- reach$B_Reach_cms - reach$R_Reach_cms

# Plot the reach differences
reach.plot <- ggplot(reach, aes(x = Diff))
reach.plot +
  geom_histogram()

# There is an outlier that makes no sense, so remove it and replot
reach <- reach[reach$Diff > -50,]
reach.plot <- ggplot(reach, aes(x = Diff))
reach.plot +
  geom_histogram(binwidth = 2)

# Remove all cases where the players had equal reach
reach <- subset(reach, !(Diff == 0))

# Identify the fighter with the longer reach
reach$Advantage <-
  case_when(
    reach$Diff > 0 ~ 'Blue',
    reach$Diff < 0 ~ 'Red'
  )

# Identify if the advantaged fighter won
reach$AdWin <-
  case_when(
    reach$Advantage == reach$Winner ~ 'Win',
    reach$Advantage != reach$Winner ~ 'Lose'
  )

reach$AdWin <- as.factor(reach$AdWin)

# Take the absolute value of the difference
reach$AbsDiff <- abs(reach$Diff)

# Plot the data
reach.hist <- ggplot(reach, aes(x = AbsDiff))
reach.hist +
  geom_histogram(binwidth = 2) +
  facet_wrap(~ AdWin, ncol = 1)

# Create the logistic regression
reach.model <- glm(AdWin ~ AbsDiff, data = reach, family = binomial())

# Display summary
summary(reach.model)

```

There is a statistically significant correlation between the winner and having a reach advantage. (Note: Need to determine how to interpret the 0.02 coefficient. Is that a 2% increase in winning percentage per 1 cm of reach? That seems like a lot.)

## Check to see how gender affects this

```{r reach_gender_analysis}

# Plot the data
reach_gender.hist <- ggplot(reach, aes(x = AbsDiff, color = gender))
reach_gender.hist +
  geom_histogram(binwidth = 2, aes(fill = gender)) +
  facet_grid(AdWin ~ gender)

# Create the logistic regression
reach_diff.model <- glm(AdWin ~ AbsDiff, data = reach, family = binomial())
reach_gender.model <- update(reach_diff.model, .~. + gender)

# Display summary
summary(reach_diff.model)
summary(reach_gender.model)
anova(reach_diff.model, reach_gender.model)


```

It appears that adding gender does not add much to the model.

# Inquiry 3: Does height have an advantage?

```{r height_analysis}
# Create a smaller data frame
height <- df[,c('Winner', 'B_Height_cms', 'R_Height_cms')]

# Calculate the difference in reach (positive = blue advantage)
height$Diff <- height$B_Height_cms - height$R_Height_cms

# Remove all cases where the players had equal reach
height <- subset(height, !(Diff == 0))

# Identify the fighter with the longer reach
height$Advantage <-
  case_when(
    height$Diff > 0 ~ 'Blue',
    height$Diff < 0 ~ 'Red'
  )

# Identify if the advantaged fighter won
height$AdWin <-
  case_when(
    height$Advantage == height$Winner ~ 'Win',
    height$Advantage != height$Winner ~ 'Lose'
  )

height$AdWin <- as.factor(height$AdWin)

# Take the absolute value of the difference
height$AbsDiff <- abs(height$Diff)

# Plot the data
height.hist <- ggplot(height, aes(x = AbsDiff))
height.hist +
  geom_histogram(binwidth = 2) +
  facet_wrap(~ AdWin, ncol = 1)



# Create the logistic regression
height.model <- glm(AdWin ~ AbsDiff, data = height, family = binomial())

# Display summary
summary(height.model)

```

There is no significant advantage.

# Inquiry 4: What about combining height and reach?

```{r height_reach_analysis_1}
# Create a smaller data frame
height_reach <- df[,c('Winner', 'B_Reach_cms', 'R_Reach_cms', 'B_Height_cms', 'R_Height_cms')]

# Calculate the difference in height and reach relative to Blue
height_reach$HDiff <- height_reach$B_Height_cms - height_reach$R_Height_cms
height_reach$RDiff <- height_reach$B_Reach_cms - height_reach$R_Reach_cms

# Draw a scatterplot of the differences
height_reach.scatter <- ggplot(height_reach, aes(x = RDiff, y = HDiff))
height_reach.scatter +
  geom_point() +
  geom_smooth(method = 'lm')

# Remove bad data point and replot
height_reach <- height_reach[abs(height_reach$RDiff) < 50,]
height_reach.scatter <- ggplot(height_reach, aes(x = RDiff, y = HDiff))
height_reach.scatter +
  geom_point() +
  geom_smooth(method = 'lm')
```

This scatterplot only shows that the height and reach advantages have a correlation with each other, but it does not say anything about wins and losses.

```{r height_reach_analysis_2}
# Calculate the mean advantages for the winner
stat.desc(cbind(height_reach$HDiff, height_reach$RDiff))

```

Blue does not appear to have any average advantage/disadvantage with respect to height and reach.

```{r height_reach_analysis_3}
# Create a column that indicates whether blue wins/loses
height_reach$BlueWin <-
  case_when(
    height_reach$Winner == 'Blue' ~ 'Yes',
    height_reach$Winner == 'Red' ~ 'No'
  )
height_reach$BlueWin <- as.factor(height_reach$BlueWin)

# Create a logistic regression
height_reach.model <- glm(BlueWin ~ HDiff * RDiff, data = height_reach, family = binomial())

summary(height_reach.model)
```

As before, height does not appear to have an impact, but reach does.

# Inquiry 5: Compare stances

```{r stance_analysis}
# Create a smaller data frame
stance <- df[,c('Winner', 'B_Stance', 'R_Stance')]

# Only keep matches where the stances are different
stance <- subset(stance, !(B_Stance == R_Stance))

```

```{r}
# Height, Weight, Age, Experience (# of matches), Wins, Stance


ggplot(df, aes(x = gender, fill = Winner)) +
  geom_bar()
```
```{r}
df %>%
  select(gender, Winner) %>%
  dplyr::summarise(`% Win` = n())
```



# Inquiry 6: Does a younger age lead to more wins?

```{r age_analysis}
# Create a smaller data frame
age <- df[,c('Winner', 'B_age', 'R_age', 'age_dif')]
# Plot the age differences
age.plot <- ggplot(age, aes(x = age_dif))
age.plot +
  geom_histogram()
# Remove all cases where the players had equal age
age <- subset(age, !(age_dif == 0))
# Identify the fighter with the younger age (advantage = blue)
age$Advantage <-
  case_when(
    age$age_dif < 0 ~ 'Blue',
    age$age_dif > 0 ~ 'Red'
  )
# Identify if the advantaged fighter won
age$AdWin <-
  case_when(
    age$Advantage == age$Winner ~ 'Win',
    age$Advantage != age$Winner ~ 'Lose'
  )
age$AdWin <- as.factor(age$AdWin)
# Take the absolute value of the difference
age$AbsDiff <- abs(age$age_dif)
# Plot the data
age.hist <- ggplot(age, aes(x = AbsDiff))
age.hist +
  geom_histogram(binwidth = 2) +
  facet_wrap(~ AdWin, ncol = 1)
# Create the logistic regression
age.model <- glm(AdWin ~ AbsDiff, data = age, family = binomial())
# Display summary
summary(age.model)
```
